---
title: "Intro to tidyverse"
author: "HBH"
date: "26/05/2021"
---

#Tidyverse: a one stop shop for data exploration
```{r}
library(tidyverse)
```


#> ── Attaching packages ─────────────────────────────────────── tidyverse 1.3.0 ──
#> ✔ ggplot2 3.3.2     ✔ purrr   0.3.4
#> ✔ tibble  3.0.3     ✔ dplyr   1.0.2
#> ✔ tidyr   1.1.2     ✔ stringr 1.4.0
#> ✔ readr   1.4.0     ✔ forcats 0.5.0

ggplot2:    Grammar for creating graphics
purrr	R:    Functional programming
tibble:	    Modern and effective table system
dplyr:	    Grammar for data manipulation
tidyr:	    Set of functions to create tidy data
stringr:	  Function set to work with characters
readr:    	An easy and fast way to import data
forcats:	  Tools to easily work with factors


##The grammar
magrittr introduces a modern grammar to R (inherited from other languages) in the form of a pipe %>% 

```{r}
mean(1:5)
```

```{r}
1:5 %>% mean()
```


##The vocabulary of dplyr
dplyr introduces a new vocabulary to the R language with a set of useful tools

```{r}
mtcars
```

###select()
```{r}
mtcars %>% select(mpg:disp)
```

###mutate()
```{r}
mtcars %>% select(mpg:cyl) %>% 
  mutate(mpg_per_cyl = mpg / cyl) 
```

###filter()
```{r}
mtcars %>% select(mpg:cyl) %>% 
  mutate(mpg_per_cyl = mpg / cyl) %>% 
  filter(cyl == 8) 
```


###arrange()
```{r}
mtcars %>% select(mpg:cyl) %>% 
  mutate(mpg_per_cyl = mpg / cyl) %>% 
  filter(cyl == 8) %>% 
  arrange(mpg_per_cyl) 
```


####Excercise 1
Which car has the lowest horsepower (hp) to weight ratio (wt)?
```{r}


```




####Shortcut your pipe! 
Tools > Modify Keyboard Shortcuts > Insert Pipe Operator (use a semi-colon)


###summarise()
Reducing your data: 

What's the average quarter mile time (qsec) of 4, 6, and 8 cylinder engine cars
In base R...
```{r}
#Simple mean
mean(mtcars$qsec)
#Mean by groups
aggregate(mtcars$qsec, list(mtcars$cyl), FUN = mean)
```
In tidyverse!!!
```{r}
#Simple mean
mtcars %>% summarise(mean = mean(qsec))
#Mean by groups
mtcars %>% group_by(cyl) %>% summarise(mean = mean(qsec))
#Add lots of arguments and functions!
mtcars %>% group_by(cyl, vs) %>% 
  summarise(mean = mean(qsec), 
            min  = min(qsec), 
            max  = max(qsec))
```

###count()
```{r}
mtcars %>% group_by(cyl, vs) %>% 
  count()
```

###join
```{r}
full_join(
  
  mtcars %>% group_by(cyl, vs) %>% 
  summarise(mean = mean(qsec), 
            min  = min(qsec), 
            max  = max(qsec))
  ,
  
  mtcars %>% group_by(cyl, vs) %>% 
  count()
  
  , by = c("cyl", "vs")
)
```




#Wrangling the tidyverse 

##operators
```{r}
mtcars %>% filter(cyl >= 6)
mtcars %>% filter(! cyl == 4)
mtcars %>% filter(cyl <= 8 & cyl > 4)
mtcars %>% filter(cyl == 6 | cyl  == 8)
mtcars %>% filter(cyl %in% c(6,8))
```
##stringr
```{r}
mtcars %>% rownames_to_column(var = "cars") %>% 
  filter(str_detect(cars, pattern = "Mazda"))
```

```{r}
mtcars %>% rownames_to_column(var = "cars") %>% 
  mutate(cars = str_replace(cars, pattern = "Mazda", replacement = "Mazzy"))
```


##forcats
```{r}
mtcars %>% mutate(cyl = as.factor(cyl))
```
```{r}
mtcars %>% mutate(cyl = as.factor(cyl)) %>% 
  pull(cyl) %>% 
  fct_count()
```
```{r}
mtcars %>% mutate(cyl = factor(cyl)) %>% 
  mutate(cyl = fct_lump(cyl, n = 1))
```

##pivots
```{r}
mtcars %>% rownames_to_column(var = "cars") %>% 
  filter(str_detect(cars, pattern = "Merc")) %>% 
  group_by(cars, cyl) %>% 
  summarise(mean = mean(qsec)) %>% 
  pivot_wider(names_from = cyl, values_from = mean)


```

#Plotting
```{r}
library(ggrepel)
mtcars %>% rownames_to_column(var = "cars") %>% 
  filter(str_detect(cars, pattern = "Merc")) %>% 
  ggplot(aes(y = mpg, x = wt)) +
  geom_point() +
  geom_text_repel(aes(label = cars))
```

```{r}
library(patchwork)

p1 = mtcars %>% rownames_to_column(var = "cars") %>% 
  filter(str_detect(cars, pattern = "Merc")) %>% 
  mutate(cyl = as.factor(cyl)) %>% 
  ggplot(aes(y = mpg, x = wt, col = cyl)) +
  geom_point() +
  geom_text_repel(aes(label = cars)) 

p2 = mtcars %>% rownames_to_column(var = "cars") %>% 
  filter(str_detect(cars, pattern = "Merc")) %>% 
  mutate(cyl = as.factor(cyl)) %>% 
  group_by(cyl) %>% 
  summarise(mean = mean(qsec),
            se = sd(qsec) / sqrt(n())) %>% 
  ggplot(aes(x = cyl, y = mean, col = cyl)) +
  geom_pointrange(aes(ymin = mean - se, ymax = mean + se))

p1 + p2
```



#Resources
https://r4ds.had.co.nz/index.html
https://www.rstudio.com/resources/cheatsheets/

