---
title: "Tidyverse Tricks"
author: "JPWR"
date: "20/10/2021"
output: 'html_document'
---

# Tidyverse: a one stop shop for data exploration

"The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures."

**Warning:** this demo jumps all over the place!

```{r}
# install.packages("tidyverse")
library(tidyverse)

## alternatively, a 1-line package loader:
pacman::p_load(tidyverse, cowplot, gridExtra)

## Note: key shortcuts for pipes %>% hallelujah (+ props to Hugo Harrison)

# Tools > Modify Keyboard Shortcuts > Insert Pipe Operator (use a semi-colon)

## load some datasets
library(gapminder)
data(gapminder)
data(starwars)
```


#> ── Attaching packages ─────────────────────────────────────── tidyverse 1.3.0 ──
#> ✔ ggplot2 3.3.2     ✔ purrr   0.3.4
#> ✔ tibble  3.0.3     ✔ dplyr   1.0.2
#> ✔ tidyr   1.1.2     ✔ stringr 1.4.0
#> ✔ readr   1.4.0     ✔ forcats 0.5.0

## forcats (https://forcats.tidyverse.org/)

Factors are very useful for plotting and very annoying to code efficiently. forcats makes everything better.

```{r}
## count factors
gapminder %>% count(continent, sort=TRUE) ## count a factor
gapminder %>% add_count(continent, sort=TRUE) ## add count to your dataframe

## reorder by the number of observations in each level
star_hair<-starwars %>% mutate(hair_color = fct_infreq(hair_color))
levels(star_hair$hair_color)

## Remember that changing factor levels has no effect on the order of your dataframe. Factor order appears in models and plots
ggplot(starwars, aes(hair_color)) + geom_bar()
ggplot(star_hair, aes(hair_color)) + geom_bar()

# Notice the NA levels and long tail of rare hair colours. fct_lump*() lumps rare levels together:

## use fct_lump_min to assign 'Other' to any levels fewer than 'min'
star_lump<-starwars %>% mutate(hair_color = fct_lump_min(hair_color, min = 2))
ggplot(star_lump, aes(hair_color)) + geom_bar()

## use fct_lump_prop to assign 'Other' to any levels less than 10% of observations
star_lump<-starwars %>% mutate(hair_color = fct_lump_prop(hair_color, prop = 0.1))
ggplot(star_lump, aes(hair_color)) + geom_bar()

## use fct_reorder to order levels by a number variable
gapminder %>% filter(continent == "Africa") %>% group_by(country) %>% summarise(lifeExp = mean(lifeExp)) %>% 
              mutate(country = fct_reorder(country, lifeExp)) %>% 
              ggplot() + geom_point(aes(country, lifeExp)) +
              coord_flip()
```

## readr (https://readr.tidyverse.org/)

Try switching over to readr functions to read your data in. They work just the same as base, but with more helpful messaging + functionality:

```{r}
u5mr<-read_csv('unicef-u5mr.csv')

## pipe to load and bind multiple csvs into 1 object
tbl <-
    list.files(path = './csv/', pattern = "*.csv", full.names=TRUE) %>% 
    map_df(~read_csv(.))
```


## filter (https://dplyr.tidyverse.org/)

Filters are quick one-liners to subset your dataset into the rows and columns you're interested in. The basic command is 'filter', but there are other filter-type verbs in tidyverse:

```{r}
## Taking a conditional filter
starwars %>% filter(mass > mean(mass, na.rm = TRUE)) ## Any character with mass above the mean mass

## Taking the top values: slice
gapminder %>% slice_max(gdpPercap, n = 10)
gapminder %>% slice_min(gdpPercap, n = 10)

## Filter and select using a string search 
gapminder %>% filter(str_detect(country, 'Rep.')) %>% distinct(country) # find the republics

## take columns that match a condition (good for wide data)
u5mr <- read.csv("unicef-u5mr.csv")
u5mr %>% select(CountryName, starts_with('U5MR.20'))

```

## Building tidy code is powerful, efficient, and easy to read

Almost everything you need in R can be done in the tidyverse. We can pipe to base R functions, and other tidyverse packages:

```{r}
starwars %>% dim()
starwars %>% droplevels()
starwars %>% ggplot() + geom_point(aes(mass, birth_year))

## fitting a linear model (https://broom.tidymodels.org/)
starwars %>%
  lm(mass ~ birth_year, data = .) %>%
  broom::tidy()

starwars %>%
  lm(mass ~ height, data = .) %>%
  broom::glance() %>% 
  write.csv(file = 'model_output.csv')
```

  
# Resources
https://forcats.tidyverse.org/index.html
https://dplyr.tidyverse.org/
https://broom.tidymodels.org/
https://monashbioinformaticsplatform.github.io/r-more/topics/tidyverse.html


