---
title: "Update R"
author: "LBE"
date: "25/05/2022"
output: html_document
---

To update R, without losing all your packages, follow theses steps:
[stolen with modifications from https://www.r-bloggers.com/2017/05/upgrade-r-without-losing-your-packages/]


#1 - Build temp file with packages
```{r}
tmp <- installed.packages()
installedpkgs <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
save(installedpkgs, file="installed_old.rda") #saves on you working directory - make sure you know where that is
```

#2 - Install new version of R

```{r}
#to figure out what version you're currently on
sessionInfo()
```

ON MAC: Download most recent version from https://cran.r-project.org 
ON PC: R on a university computer should be installed from AppsAnywhere, not from CRAN. 

From AppsAnywhere: "Due to issues with working from network drives for some features of R, on University machines a temporary location on the computers local storage will be created on launch, with a shortcut on the desktop for access. This will be the working directory when launching R Studio and so all files should be put here. On exit this data will backed up to a folder in your H: drive called 'r_working' within 'Documents'. Furthermore, if this location exists on your H: drive from a previous launch it will also copy the contents to the local storage ready for use again on subsequent launches."


**Remember that updating RStudio is NOT the same thing as updating R! (You should probably do both though...)**

#3 - Reload packages
```{r}
tmp <- installed.packages()
load("installed_old.rda")
installedpkgs.new <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
missing <- setdiff(installedpkgs, installedpkgs.new)
install.packages(missing) #WILL TAKE SEVERAL MINUTES (9min)
update.packages(ask = FALSE) #CAN TAKE EVEN LONGER (17min)
```

#Bonus: New packages?

*For PC `installr`
*For Mac `UpdateR` [but notes that the update package function doesn't operate well outside of base R app.]

#4 - Bioconductor packages
if you use bioconductor, this will apparently work. I haven't tested it so feel free to edit if you use it and do any changes :)

```{r}
chooseBioCmirror()
biocLite() 
load("installed_old.rda")
tmp <- installed.packages()
installedpkgs.new <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
missing <- setdiff(installedpkgs, installedpkgs.new)
for (i in 1:length(missing)) biocLite(missing[i])
```

