---
title: "Intro to ggplot"
author: "LBE"
date: "09/02/2022"
output: html_document
---

# Welcome to ggplot!
* Based on tidyverse principles
* Iterative principles, layering data and graphical themes
* Uses a set Grammar of Graphics:
* **Data** is **mapped** to aesthetics
   - **Layer**s are **geom**etric elements, and **stat**istical transformations
   - **scale**s map values to geometric space
   - **coord**inate systems decsribed how data coordinates are mapped to a graphical plane
   - **Facet**s breaks up and displays subsets of data
   - **Theme**s control details about the display, like font size etc. 

# Libraries
```{r, message=FALSE} 
library(tidyverse)
library(patchwork)
library(ggpubr)
``` 

# Data
Today's data: a collection of dispersal and self-recruitment data for anemonefish
```{r}
load("~/Documents/Work/R-Scripts/use-r/demos/ggplot/NemoDatabase.Rdata")

```   


Let's start with some data wrangling
```{r}
nemo_dat = nemo %>% 
  mutate(Species = paste("A.", Species),
         Range_disp = as.character(Range_disp),
         PLD = as.character(PLD),
         Mkr_n = as.character(Mkr_n)) %>% 
  mutate(SR_mean = ifelse(is.na(SR_mean), (SR_min + SR_max) / 2, SR_mean)) %>% 
  mutate(Method = str_replace(Method, "IBD", "Genetic structure"),
         Method = str_replace(Method, "Assignment", "Assignment tests"),
         Method = factor(Method, levels = c("Parentage", "Assignment tests", "Genetic structure", "Otolith tagging"))) %>% 
  mutate(Range_disp.1 = as.character(Range_disp)) %>% 
  separate(Range_disp.1, c("min_Range", "max_Range"), sep = "-") %>% 
  mutate(min_Range = as.numeric(min_Range),
         max_Range = as.numeric(max_Range)) %>% 
  mutate(Adults = replace_na(Adults, 0),
         Juveniles = replace_na(Juveniles, 0)) %>% 
  mutate(Specis = as.factor(Species))
nemo_dat
```

# Fig 1. 
Our first figure!

Every ggplot pot has three main components:

* **data**
* **aes**thetic mappings between data & visual properties
* At least one **layer** that describes how each observation is rendered, usually a **geom** functions

```{r}
ggplot(data = nemo_dat, aes(x = Publication_year, fill = Method)) +
  geom_histogram()
```

## p1
```{r}
p1 <- ggplot(nemo_dat, aes(x = Publication_year, fill = Method)) +
  geom_histogram(binwidth = 1, col = "white", size = 1) +
  scale_fill_manual(values = c("#4e79a7", "#26A69A", "#edc948",  "#f28e2b")) +
  labs(x = "Publication year", y = "Number of studies") +
  scale_y_continuous(expand = c(0,0), limits = c(0,5.5), breaks = c(0,2,4,6)) #
p1
```

```{r}
p1 = p1 +
  theme_classic()
p1
```

```{r}
p1 +
  theme_classic() + 
  theme(axis.line = element_blank(),
        panel.border = element_rect(size = 1, fill = "transparent"),
        legend.title = element_blank(),
        legend.position = c(.15,.8),
        plot.margin = margin(.2,.1,.1,.4, unit = "cm"))
```

```{r}
Fig2_theme =  theme_classic() + 
  theme(axis.line = element_blank(),
        panel.border = element_rect(size = 1, fill = "transparent"),
        legend.title = element_blank(),
        legend.position =  "none",
        legend.background = element_rect(fill = "transparent"),
        plot.margin = margin(.2,.1,.1,.4, unit = "cm")) 
### add elements

p1 = p1 + Fig2_theme
p1
```

## p2
```{r}
my_species = c("A. percula", "A. polymnus", "A. chrysopterus", "A. clarkii", 
  "A. perideraion", "A. bicinctus", "A. frenatus", "A. melanopus", "A. ocellaris", "A. latezonatus",  
  "A. omanensis", "A. mccullochi", "A. akallopisos", "A. sandaracinos")

p2 = ggplot(nemo_dat %>% mutate(Species = factor(Species, levels = rev(my_species))), 
       aes(x = Species, fill = Method)) +
  geom_histogram(stat = "count", width = 1, col = "white", size = 1) +
  scale_fill_manual(values = c("#4e79a7", "#26A69A", "#edc948",  "#f28e2b")) +
  labs(x = "", y = "Number of studies") +
  scale_y_continuous(expand = c(0,0), breaks = c(2,4,6,8), limits = c(0, 10)) +
  coord_flip() + ## flip it!
  Fig2_theme +
  theme(axis.text.y = element_text(face = "italic"), #edit axis
        legend.position = c(.6,.2))

p2
```

## p3
```{r}
p3 = ggplot(nemo_dat, aes(x = Time_ya, fill = Method)) +
  geom_histogram(binwidth = 1, col = "white", size = 1) +
  labs(x = "Study duration", y = "Number of studies") +
  scale_fill_manual(values = c("#4e79a7", "#26A69A", "#edc948",  "#f28e2b")) +
  scale_y_continuous(expand = c(0,0), limits = c(0,14), breaks = c(0,3,6,9,12)) +
  scale_x_continuous(breaks = c(0,2,4,6,8,10)) +
  Fig2_theme

p3
```

## patchwork
```{r}
p1 + p2 
p1/p3
p2 + p1/p3
```



## ggpubr
```{r}
ggarrange(p1,p2, align = "h")
ggarrange(p1,p3, nrow = 2, align = "v")
ggarrange(p2, 
          ggarrange(p1,p3, nrow = 2), ncol = 2)
```

Shared legends
```{r}
ggarrange(p2, 
          ggarrange(p1,p3, nrow = 2, legend = "none"), ncol = 2, common.legend = TRUE, legend = "top" )
```



Panel labels
```{r}
right.panel = ggarrange(p1,p3, nrow = 2, legend = "none", 
                        labels = c("b", "c"), 
                        font.label = list(size = 10, color = "black", face = "bold", family = "Helvetica", position = "top"))

ggarrange(p2, right.panel, ncol = 2,
          common.legend = TRUE, legend = "bottom",
          labels = c("a"), 
          font.label = list(size = 10, color = "black", face = "bold", family = "Helvetica", position = "top"))
```

Widths and heights
```{r}
Fig2 = ggarrange(p2, right.panel, ncol = 2, widths = c(.7,1), 
                 common.legend = TRUE, legend = "bottom",
                 labels = c("a"), 
                 font.label = list(size = 10, color = "black", face = "bold", family = "Helvetica", position = "top"))
Fig2
```

I love it! 
```{r}
ggsave("./PublicationFigures/figures/Fig2.1.pdf", plot = Fig2, width = 174, height = 100, units = "mm", dpi = 600)
```


# Fig 3. Complex figures
## a. Insets

```{r}
p3 + annotation_custom(
    ggplotGrob(p1), 
    xmin = 2, xmax = 11, ymin = 7, ymax = 13.5)
```

## b. annotations
```{r}  
p3 + annotate("segment", x = 0, xend = 3, y = 9, yend = 5, colour = "black", size = 1) +
  annotate("text", x = 3.5, y = 5, label = "Most studies are short", 
           colour = "black", size = 5, fontface = 2, family = "Helvetica", hjust = 0)
```

## c. difficult text
```{r}
#text
p1 + labs(x = expression(lots~of~text))
#Symbols
p1 + labs(x = expression("SR" %+-% "1sd" ))
#superscript
p1 + labs(x = expression(ha.m^2))
#multiple lines
p1 + labs(x = expression(atop(Two~lines, mean^super %+-% "1sd")))
```

## d. text justification
```{r}
td <- expand.grid(
    hjust=c(0, 0.5, 1),
    vjust=c(0, 0.5, 1),
    angle=c(0, 45, 90),
    text="text"
)

ggplot(td, aes(x=hjust, y=vjust)) + 
    geom_point() +
    geom_text(aes(label=text, angle=angle, hjust=hjust, vjust=vjust)) + 
    facet_grid(~angle) +
    scale_x_continuous(breaks=c(0, 0.5, 1), expand=c(0, 0.2)) +
    scale_y_continuous(breaks=c(0, 0.5, 1), expand=c(0, 0.2))
```


## e. difficult legends
```{r}
#Edit the legend
p3.legend = p3 +
  theme(legend.background = element_rect(fill = "transparent"),
        legend.text = element_text(face = "italic", size = 10),
        legend.title = element_blank(),
        legend.position = c(.6,.8),
        legend.spacing.x = unit(.5, 'cm'),
        legend.spacing.y = unit(.5, 'cm'),
        legend.key = element_rect(colour = "white",size = 2),
        legend.key.size = unit(.6, "cm")) +
  guides(fill=guide_legend(nrow=2, byrow=FALSE))

p3.legend

#Extract the legend
legend = as_ggplot(get_legend(p3.legend))

#plot the legend like a plot 
ggarrange(p3, legend, nrow = 2)
```

## f. facetting
```{r}
p3 + facet_wrap(~Method) +
  theme(strip.background = element_blank(),
        strip.text = element_blank()) +
  geom_text(data = data.frame(Method = c("Parentage", "Assignment tests", "Genetic structure", "Otolith tagging")), 
            aes(x = 10, y = 12, label = Method), hjust = 1)
```

```{r}
data("ToothGrowth")

ToothGrowth = ToothGrowth %>% 
  mutate(dose = as.factor(dose))
p <- ggplot(ToothGrowth, aes(x = dose, y = len, fill = dose)) +
  geom_boxplot() +
  theme_classic()
p
 
# Add p-values comparing groups
# Specify the comparisons you want
my_comparisons <- list( c("0.5", "1"), c("1", "2"), c("0.5", "2") )
p + stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 50)                   # Add global p-value 
```

# Playing with colour
*Personalised colour palettes
*library(wesanderson)
*https://coolors.co/
*https://medialab.github.io/iwanthue/



