---
title: "Demo on Mapping in R"
author: "Eva Maire - LEC"
date: "17/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is a demo on mapping in base R and using ggplot2. 
If you want to replicate this demo please download the csv file "cities_UK.csv" and update the 'path' below. 

## Example in base R 

```{r , echo=T}
path <- "/Volumes/EM2T/Lancaster/LEC_coding_group/mapping"
setwd(path)
```

Load libraries
```{r , echo=T}
library(plot3D)
library(maps)
```

Load dataset
```{r , echo=T}
cities <- read.csv("cities_UK.csv")
head(cities)
```

Create a new variable that will be used later:
```{r , echo=T}
cities$logPop <- log(cities$POP)
```

Check the range of lat and long values to define the extent of the map
```{r , echo=T}
summary(cities[,c("Longitude","Latitude")]) 
lon <- c(-6, 0)
lat <-  c(50, 57)
```

Draw a map using scatter2D and map:  
Remember that in base R: the first plot is displayed and then, other elements can be added with add=TRUE.   
If add=FALSE, another plot is created and displayed

Initial plot: point are not drawn (cex = 0) but this sets up the spatial extent
```{r , echo=T, fig.height = 7, fig.width = 7, fig.align = "center"}
scatter2D(lon, lat,xlab="Lon", ylab="Lat", cex=0) # Points are not drawn (cex = 0) but sets up the spatial extent
```

Then, add base map:
```{r , echo=T, fig.height = 7, fig.width = 7, fig.align = "center"}
scatter2D(lon, lat,xlab="Lon", ylab="Lat", cex=0) # Points are not drawn (cex = 0) but sets up the spatial extent
maps::map(database="world",border=NA,fill=TRUE,col="grey80",add=TRUE) 
```

Display cities:
```{r , echo=T, fig.height = 7, fig.width = 7, fig.align = "center"}
scatter2D(lon, lat,xlab="Lon", ylab="Lat", cex=0) # Points are not drawn (cex = 0) but sets up the spatial extent
maps::map(database="world",border=NA,fill=TRUE,col="grey80",add=TRUE) # add base map
scatter2D(cities$Longitude, cities$Latitude,
          cex=1, pch=19, col = "black", add=TRUE) # cex = 1 and points are drawn this time

```

Add names and title:
```{r , echo=T, fig.height = 7, fig.width = 7, fig.align = "center"}
scatter2D(lon, lat,xlab="Lon", ylab="Lat", cex=0) # Points are not drawn (cex = 0) but sets up the spatial extent
maps::map(database="world",border=NA,fill=TRUE,col="grey80",add=TRUE) # add base map
scatter2D(cities$Longitude, cities$Latitude,
          cex=1, pch=19, col = "black", add=TRUE) #Points are drawn this time
text2D(cities$Longitude, cities$Latitude-0.1,labels=cities$CITY_NAME, add=T) #Allows to add text based on coordinates
title("Major Cities in the UK")
```


Display cities based on LOG population counts:
```{r , echo=T, fig.height = 7, fig.width = 7, fig.align = "center"}
par(oma=c(1,1,1,1)) #increase margins to draw a legend next to the map

scatter2D(lon, lat,xlab="Lon", ylab="Lat", cex=0) # Points are not drawn (cex = 0) but sets up the spatial extent
maps::map(database="world",border=NA,fill=TRUE,col="grey80",add=TRUE) # add base map
scatter2D(cities$Longitude, cities$Latitude,
          cex=2,pch=19,
          colvar = cities$logPop,add=TRUE,
          colkey = list(side = 4, 
                        dist = 0.07)) # colkey is a list of parameters for the legend
text2D(cities$Longitude, cities$Latitude-0.1,labels=cities$CITY_NAME, add=T)
title("Major Cities in the UK (LOG Population)")
```

```{r , echo=F}
graphics.off()
rm(list=ls())
```


## Example with ggplot2

We are going to draw the same map using ggplot2 (and other libraries) this time.  
```{r , echo=T}
#Load library
library(ggplot2)
```

ggplot2 offers global base map which can be customized

Basic World Map:
```{r , echo=T}
world_map <- map_data("world") #ggplot2
ggplot(world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill="lightgray", colour = "white")
```

Major Cities in the UK:
```{r , echo=T}
cities <- read.csv("cities_UK.csv")
cities$logPop <- log(cities$POP)
summary(cities[,c("Longitude","Latitude")]) #Check the range of lat and long values to define the extent of the map

uk <- ggplot() + 
  geom_polygon(data = world_map, aes(x=long, y = lat, group = group),fill = "grey", color = "grey") +
  coord_fixed(xlim = c(-6, 0),  ylim = c(50, 57), ratio = 1.3)+ # x = long / y = lat
  geom_point(data = cities,aes(x=Longitude, y=Latitude, fill = logPop), colour="black", size=5, pch=21)+
  scale_fill_continuous(name = "LOG Population") +
  scale_x_continuous("longitude")+
  scale_y_continuous("latitude")+  
  theme_classic() + theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))

uk

```

Add city names with 'ggrepel' package to avoid overlapping labels:
```{r , echo=T}
library(ggrepel) 

uk_name <- ggplot() + 
  geom_polygon(data = world_map, aes(x=long, y = lat, group = group),fill = "grey", color = "grey") +
  coord_fixed(xlim = c(-6, 0),  ylim = c(50, 57), ratio = 1.3)+ # x = long / y = lat
  geom_point(data = cities,aes(x=Longitude, y=Latitude, fill = logPop), colour="black", size=5, pch=21)+
  geom_text_repel(data = cities , aes(x=Longitude, y=Latitude,label=CITY_NAME), direction = "both",size=3,force = 5)+ 
  scale_fill_continuous(name = "LOG Population") +
  scale_x_continuous("longitude")+
  scale_y_continuous("latitude")+  
  theme_classic() + theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))

uk_name
```

Change color scale with RColorBrewer:
```{r , echo=T}
library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(4, "Spectral"))) # Use 'rev' to reverse the original color palette (red > yellow > blue)

uk_color <- ggplot() + 
  geom_polygon(data = world_map, aes(x=long, y = lat, group = group),fill = "grey", color = "grey") +
  coord_fixed(xlim = c(-6, 0),  ylim = c(50, 57), ratio = 1.3)+ # x = long / y = lat
  geom_point(data = cities,aes(x=Longitude, y=Latitude, fill = logPop), colour="black", size=5, pch=21)+
  geom_text_repel(data = cities , aes(x=Longitude, y=Latitude,label=CITY_NAME), direction = "both",size=3,force = 5)+ 
  scale_fill_gradientn(name = "LOG Population", colours = myPalette(100),limits=c(min(cities$logPop),max(cities$logPop))) +
  scale_x_continuous("longitude")+
  scale_y_continuous("latitude")+  
  theme_classic() + theme(panel.border = element_rect(colour = "black", fill=NA, size=1.5))

uk_color
```

Geocoding with ggmap and Google API in R => check https://evamaire.com/2020/04/16/geocoding-with-ggmap-in-r/

Thanks!